---
title: "Pipeline Comparisons"
author: "Amanda D. Clark"
output: html_notebook
---
# Purpose
WORDS
## Sources & Resources 
Sources and resources are linked where applicable
https://rdrr.io/cran/FactoMineR/man/coeffRV.html
https://cran.r-project.org/web/packages/UpSetR/vignettes/basic.usage.html

# Setting Up
## Setting Up Environment 

```{r}
# clear workspace
rm(list=ls(all.names=TRUE))

# Function to check for a package host on CRAN, then install (if needed) and library the package
prep_cranpack <- function (x){
if (!requireNamespace(x, quietly = TRUE)) {
  install.packages(x)
  library(x, character.only = TRUE, quietly = TRUE)
} else {
  library(x, character.only = TRUE, quietly = TRUE)
}}

# Function to check for a package host on bioconductor, then install (if needed) and library the package
prep_biocpack <- function (x){
if (!requireNamespace(x, quietly = TRUE)) {
  BiocManager::install(x)
  library(x, character.only = TRUE, quietly = TRUE)
} else {
  library(x, character.only = TRUE, quietly = TRUE)
}}

# loading list of CRAN packages
cranpacks <- c("BiocManager", "tools", "devtools", "tidyverse", "RColorBrewer", "stringr", "UpSetR", "FactoMineR", "corrplot")
invisible(lapply(cranpacks, prep_cranpack))

# loading list of Bioconductor packages
#biocpacks <- c("ballgown", "RNAseq123", "fgsea")
#invisible(lapply(biocpacks, prep_biocpack))

# directory for input files
indir <- "../R_outputs/DGE_Analyses"

# make a directory for output files
if (! dir.exists("../R_outputs/Compare_Pipelines")) {
 dir.create("../R_outputs/Compare_Pipelines")
}
outdir <- "../R_outputs/Compare_Pipelines"

```

## Setting Up Data
Wonder if i should look at filtering effects holding pipeline stagnant and then across pipelines holding filtering stagnant? Need to get DEG across pipelines.In addition, I would like to compare these pipelines at several other levels.

I have the upset plot that will allow me to observe the intersection of gene sets across different categories. I also wanted to look at the matrix correlations using RV coefficients (which can be bootstrapped for confidence estimates). I think these two ways to look at the data will tell me the same things, but one is more quantitative. 

### Mapping Data
```{r}



```


### DEG data

```{r}
# Empty list
files <- list()
# Populate list with file names; including paths
files <- list.files(path = indir, pattern = ".csv", full.names = T)

# Make list containing data from each file in files
count_data <- list(f_name = c(file_path_sans_ext(basename(files))), f_content = files %>% map(read.csv, header = T)) # populate list with file names, paths, and content
names(count_data$f_content) <- count_data$f_name # name matrices based on file names

```

## Upset Plots

### DEG 
```{r}
## Plots with hard filtered data
# Empty List
hardflist <- list()
# Add gene IDs from each of the hard filtered pipelines
for (i in 1:18) {
  hardflist <- c(hardflist, c(count_data$f_content[i][[1]][1]))
}
# Grab Relevant List Names
hardfnames <- count_data$f_name[1:18]
# Add List Names
names(hardflist) <- hardfnames

upset(fromList(hardflist), mainbar.y.label = "Intersections", sets.x.label = "# of Genes", order.by = "freq", empty.intersections = "on", nsets = 14, group.by = "sets")


# Plots with soft filtered data

# Plots with pipeline filtered data


```

